---
title: "R Notebook"
output: github_notebook
---

```{r}
library(tidyverse)
library(dslabs)
```
```{r}
path <- system.file("extdata", package = "dslabs")
print(path)
filename <- file.path(path, "fertility-two-countries-example.csv")
wide_data <- read_csv(filename)
```
```{r}
view(wide_data)
wide_data %>% gather(year, fertility, '1960':'2015')
head(wide_data)
new_tidy_data <- wide_data %>% gather(year, fertility, -country, convert = TRUE)
head(new_tidy_data)
```


```{r}
new_wide_data <- new_tidy_data %>% spread(year, fertility)
select(new_wide_data, country, '1960':'1967')
```
```{r}
filename_ <- file.path(path, "life-expectancy-and-fertility-two-countries-example.csv")
raw_data <- read_csv(filename_)
select(raw_data, 1:6)
```


```{r}
dat <- raw_data %>% gather(key, value, -country)
head(dat)
dat %>% separate(key, c('year', 'variable_name'),'_')

dat %>% separate(key, c('year','first_variable_name', 'second_variable_name'), fill = 'right')

dat %>% separate(key, c('year', 'variable_name'), extra = 'merge')
dat %>% separate(key, c('year', 'variable_name'), extra = 'merge') %>% spread(variable_name, value)

```

## Joinig Table

```{r}
library(tidyselect)
library(dslabs)
data(murders)
head(murders)
```
```{r}
data("polls_us_election_2016")
head(results_us_election_2016)
```
## Join

```{r}
library(dplyr)
new_tab <- left_join(murders, results_us_election_2016, by = "state") %>% select(-others)

head(new_tab)
```
In our case every row

```{r}
tab1 <- slice(murders, 1:6) %>% select(state, population)
tab2 <- results_us_election_2016 %>% filter(state %in%
c("Alabama","Alaska","Arizona", "California", "Connecticut", "Delaware")) %>% select(state, electoral_votes)
```
```{r}
left_join(tab1, tab2, by = "state")
```

```{r}
right_join(tab1, tab2, by = "state")
```
```{r}
inner_join(tab1, tab2 , by = "state")
```

```{r}
full_join(tab1, tab2, by = "state")
```

## Binding

we have to sort of binding. Bind by columns or by rows

####    Binding by columns
```{r}
bind_cols(a = 1:9, b = 4:12)

```

## Set operator

the are useful to manipulate data frame 

```{r}
a <- 1:20
b <- 13:30
intersect(a,b)
```

```{r}
union(a, b)
```

```{r}
tab1 <- new_tab[1:5, ]
tab2 <- new_tab[3:7,] 
setdiff(tab1, tab2)
```

```{r}
install.packages("Lahman")
library(Lahman)
data("Batting")
head(Batting)
```
This database contain data relate to baseball team. it contain summary statistic about how players performed on offense and defense for several year. The Batting data frame contain the offensive statistic for all players for many year

```{r}
Batting %>% filter(yearID == "2016") %>% arrange(desc(HR)) %>% slice(1:10)
```


## Web Scrapping


```{r}
url <- "https://web.archive.org/web/20181024132313/http://www.stevetheump.com/Payrolls.htm"
library(rvest)
h <- read_html(url)
```

```{r}
class(h)
nodes <- h %>% html_nodes("table")
node <- nodes[[1]]
```



```{r}
tab <- nodes[[8]] %>% html_table()
tab

```
```{r}
names(tab) <- c("Index", "Team", "Payroll", "Average")

```
```{r}
pay1 <- nodes[[1]] %>% html_table()
pay2 <- nodes[[2]] %>% html_table()
pay3 <- nodes[[3]] %>% html_table()
pay4 <- nodes[[4]] %>% html_table()
```

```{r}
pay4
```
```{r}
    nodes[[22]] %>% html_table()

```

## String Processing

```{r}
library(stringr)
data("reported_heights")
class(reported_heights$height)
```

```{r}
x <- as.numeric(reported_heights$height)
```

```{r}
head(x)
```
```{r}
reported_heights %>% mutate(new_height = as.numeric(height)) %>% filter(is.na(new_height)) %>% head(12)
```

```{r}
not_inches <- function(x, smallest = 50, tallest = 84){
  inches <- suppressWarnings(as.numeric(x))
  ind <- is.na(inches) |inches < smallest | inches > tallest
  ind
}

problems <- reported_heights %>% filter(not_inches(height)) %>% pull(height)
problems
```

```{r}
yes <- c("180 cm", "70 inches")
no <- c("180", "70''")
s <- c(yes, no)
str_detect(s, "cm|inches")
```

```{r}
yes <- c("5", "6", "5'10", "5 feet", "4'11")
no <- c(" ", ".", "Five", "Six")
s <- c(yes, no)
pattern <- "\\d"
str_detect(s,pattern = pattern )
```

```{r}
install.packages("htmlwidgets") 
library(htmlwidgets)
str_view(s, pattern)
```
```{r}
str_view_all(s, pattern)
```

## Character class
 [] in R
 
```{r}
str_view(s, "[56]")
```
## Anchor

```{r}
pattern <- "^\\d$"
yes <- c("1", "5", "9")
no <- c("12", "123", " 1", "a4", "b") 
s <- c(yes, no)
str_view_all(s, pattern)
```
## Quantifier

```{r}
pattern <- "^\\d{1,2}$"
yes <- c("1", "5", "9", "12")
no <- c("123", "a4", "b") 
str_view(c(yes, no), pattern)
```
```{r}
pattern <- "^[4-7]'\\d{1,2}\"$"
yes <- c("5'7\"", "6'2\"", "5'12\"")
no <- c("6,2\"", "6.2\"","I am 5'11\"", "3'2\"", "64")
s <- c(yes, no)
str_detect(s, pattern)
```

```{r}
pattern_2 <- "^[4-7]'\\s\\d{1,2}\"$"
str_subset(problems, pattern_2)
```

##  Quantifier: *, +, ?

```{r}
yes <- c("AB", "A1B", "A11B", "A111B", "A1111B") 
no <- c("A2B", "A21B")
s <- c(yes,no)
str_detect(s, "A1*B")
```

## Group

```{r}
pattern_with_groups <-  "^([4-7]),(\\d*)$"
yes <- c("5,9", "5,11", "6,", "6,1")
no <- c("5'9", ",", "2,8", "6.1.1")
s <- c(yes, no)
str_detect(s, pattern_with_groups)
```

```{r}
str_match(s, pattern_with_groups)
```

## Search and replace with Regular expression

```{r}
pattern <- "^[4-7]'\\d{1,2}\"$"
sum(str_detect(problems, pattern))
```

```{r}
problems[c(2, 10, 11, 12, 15)] %>% str_view(pattern)
```
```{r}
str_subset(problems, "inches")
```

```{r}
str_subset(problems, "''")
```

```{r}
pattern <- "^[4-7]'\\d{1,2}$"
problems %>% str_replace("feet|ft|foot", "'") %>% str_replace("inches|in|''|\"", "") %>% str_detect(pattern) %>% sum()
```

```{r}
pattern <- "^[4-7]\\s*'\\s*\\d{1,2}$"
problems %>% str_replace("feet|ft|foot", "'") %>% str_replace("inches|in|''|\"", "") %>% str_detect(pattern) %>% sum()
```

```{r}
pattern_with_groups <- "^([4-7]),(\\d*)$"
yes <- c("5,9", "5,11", "6,", "6,1")
no <- c("5'9", ",", "2,8", "6.1.1")
s <- c(yes, no)
str_replace(s, pattern_with_groups, "\\1'\\2")
```

```{r}
pattern_with_groups <-"^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$"
str_subset(problems, pattern_with_groups) %>% head()
```
```{r}
str_subset(problems, pattern_with_groups) %>%  str_replace( pattern_with_groups,  "\\1'\\2" ) %>% head()
```

```{r}
not_inches_or_cm <- function(x, smallest = 50, tallest = 84){
    inches <- suppressWarnings(as.numeric(x))
    ind <- !is.na(inches) & ((inches >= smallest & inches <= tallest) | (inches/2.54 >= smallest & inches/2.54 <= tallest))
    !ind
}

problems <- reported_heights %>% filter(not_inches_or_cm(height)) %>% pull(height)
length(problems)
```

```{r}
converted <- problems %>% str_replace("feet|foot|ft", "'") %>% str_replace("inches|in|''|\"", "") %>% str_replace("^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$", "\\1'\\2")

pattern <- "^[4-7]\\s*'\\s*\\d{1,2}$"
index <- str_detect(converted, pattern)
mean(index)
```

```{r}
converted[index]
```

```{r}
converted[!index]
```
```{r}
yes <- c("5", "6", "5")
no <- c("5'", "5''", "5'4")
s <- c(yes, no)
str_replace(s, "^([4-7])$", "\\1'0")
```

```{r}
pattern <- "^[4-7]\\s*'\\s*(\\d+\\.?\\d*)$"
yes <- c("1,7", "1, 8", "2, " )
no <- c("5,8", "5,3,2", "1.7")
s <- c(yes, no)

str_replace(s, "^([12])\\s*,\\s*(\\d*)$", "\\1\\.\\2")
```

## Triming

```{r}
s <- "Hi "
print(cat(s))
identical(s, "Hi")
```

```{r}
convert_format <- function(s){
  s %>% str_replace("feet|feet|ft", "'") %>%
    str_replace_all("inches|in|''|\"|cm|and", "") %>%   str_replace("^([4-7])\\s*[,\\.\\s+]\\s*(\\d*)$", "\\1'\\2") %>% str_replace("^([56])'?$", "\\1'0") %>% str_replace("^([12])\\s*,\\s*(\\d*)$","\\1\\.\\2") %>% str_trim()
}
```

```{r}
word_to_numbers <- function(s){
  str_to_lower(s) %>%  str_replace_all("zero", "0") %>%
    str_replace_all("one", "1") %>%
    str_replace_all("two", "2") %>%
    str_replace_all("three", "3") %>%
    str_replace_all("four", "4") %>%
    str_replace_all("five", "5") %>%
    str_replace_all("six", "6") %>%
    str_replace_all("seven", "7") %>%
    str_replace_all("eight", "8") %>%
    str_replace_all("nine", "9") %>%
    str_replace_all("ten", "10") %>%
    str_replace_all("eleven", "11")
}
```

```{r}
converted <- problems %>% convert_format() %>% word_to_numbers()
remaining_problems <- converted[not_inches_or_cm(converted)]
pattern <- "^[4-7]\\s*'\\s*\\d+\\.?\\d*$"
index <- str_detect(remaining_problems, pattern)
remaining_problems[!index]
```
```{r}
library(tidyr)
s <- c("5'10", "6'1")
tab <- data.frame(x = s)
```
```{r}
tab %>% separate(x, c("feet", "inches"), sep = "'")
```

```{r}
tab %>% extract(x, c("feet", "inches"), regex = "(\\d)'(\\d{1,2})")
```

```{r}
s <- c("5'10", "6'1\"", "5'8inches")
tab <- data.frame(x = s)
tab %>% separate(x, c("feet", "inches"), sep = "'", fill = "right")
```

```{r}
tab %>% extract(x, c("feet", "inches"), regex = "(\\d)'(\\d{1,2})")
```
```{r}
smallest <- 50
tallest <- 84
pattern <- "^([4-7])\\s*'\\s*(\\d+\\.?\\d*)$"
new_height <- reported_heights %>% mutate(original = height, height = word_to_numbers(height) %>% convert_format()) %>% extract(height, c("feet", "inches"), regex = pattern, remove = FALSE) %>% mutate_at(c("height", "feet", "inches"), as.numeric) %>% mutate(guess = 12*feet + inches) %>% mutate(height = case_when(
  !is.na(height) & between(height, smallest, tallest) ~ height,
  !is.na(height) & between(height/2.54, smallest, tallest) ~ height/2.54,
  !is.na(height) & between(height*100/2.54, smallest, tallest) ~ height*100/2.54,
  !is.na(guess) & inches < 12 & between(guess, smallest, tallest) ~ guess,
  TRUE ~ as.numeric(NA))) %>% select(-guess)
```

